\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock

### 2023-04-28 jstatsoft/article.Rnw
Added version Marianne

%% - \Abstract{} almost as usual
\Abstract{
Animals' welfare is becoming increasingly attractive, especially in livestock farming. One method to quantify an animal's welfare is measuring its rhythmicity from activities acquired with sensors (e.g., accelerometers) mounted on the animal. The Degree of functional coupling (DFC) is one of these rhythmicity indicators. Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks. Despite the importance of this indicator and its use in the literature, no open source library is available for computing it. In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides off-the-shelves functionalities for measuring animal's rhythmicity, and other valuable add-ons such as creating actrograms, computing and visualizing average activities, and the calculation of the diurnality index. We gave special attention to integrating the package in the research workflow, from raw data to creating publication-ready content.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}

@@

  Christina Umst\"atter\\
  Johann Heinrich von Th\"unen Institut f\"ur Agrartechnologie \\
  Bundesallee 47 \\
  38116 Braunschweig\\
  Deutschland\\
  E-mail: \email{christina.umstaetter@thuenen.de}\\
  URL: \url{https://www.thuenen.de/de/fachinstitute/agrartechnologie/personal/leitung/prof-dr-christina-umstaetter}\\

@@

\section[Introduction: Degree of functional coupling and animal welfare ]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}

@@

The activityis in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP. In the time domain, we can see the magnitude of the activity at a particular time stamp. In the frequency domain we can see the magnitude (power) of oscillations of specific frequencies present in the activity. For example, a  sine wave of 24 hours (Figure \ref{fig:signalVslsp}, left, first row) period could be represented with a one-peak periodogram (Figure \ref{fig:signalVslsp}, right, first row); this peak represents the frequency of a sine wave of 24 hours period. The same idea could be illustrated for a sine wave of 6 hours period (Figure \ref{fig:signalVslsp}, second row), and so on. Moreover, an activitycomposed of the addition of 24 hours and 6 hours sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row). Consequently, a periodogram is a tool that allows us to infer the power of each possible sine wave present in the activity.

>>>>

%% - \Abstract{} almost as usual # comr: changed abstract throughout.
\Abstract{
The welfare of farm animals is a growing concern to the public, yet indicators that allow to reliably quantify the welfare of these animals are scarce. One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers). The Degree of Functional Coupling (DFC) is one of these rhythmicity indicators. Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks. In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data. Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package. The package is customized for integration in the research workflow, from raw data to creating publication-ready content and therefore allows the simple application of these sophisticated methods.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Lomb-Scargle periodogram, Degree of Functional Coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}

@@

 Marianne Cockburn\\
  Agroscope, Equines \\
  Swiss National Stud \\
  Les Longs-Prés\\
  1580, Avenches \\
  Switzerland\\
  E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
  URL: \url{}\\
  \newline

@@

\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}

@@

The activity is in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP. In the time domain, we can see the magnitude of the activity at a particular time stamp. In the frequency domain we can see the magnitude (power) of oscillations of specific frequencies present in the activity. For example, a  sine wave of 24 hours (Figure \ref{fig:signalVslsp}, left, first row) period could be represented with a one-peak periodogram (Figure \ref{fig:signalVslsp}, right, first row); this peak represents the frequency of a sine wave of 24 hours period. The same idea could be illustrated for a sine wave of 6 hours period (Figure \ref{fig:signalVslsp}, second row), and so on. Moreover, an activitycomposed of the addition of 24 hours and 6 hours sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row). Consequently, a periodogram is a tool that allows us to infer the power of each possible sine wave present in the activity.
